Unifying Gompertzian growth with the communicable disease spreading paradigm

### 2023-04-04 arxiv/main.tex
respond to comments2

\author{Matz A. Haugen, \\
Independent Reseacher, Oslo\\
\texttt{matzhaugen@gmail.com} \\
  \And
  Dorothea Gilbert \\
  Independent Reseacher, Oslo\\
}

@@

A line of reasoning employed by Rypdal and Rypdal (2020) \cite{Rypdal2020} to obtain the Gompertz curve is to linearize the SIR model by assuming
both the number of infected, $I(t)$, and cumulative infected, $Y(t)$, are much less than the initial pool of susceptible people, $N \gg Y \ge I$, which seems well founded at the national level, viz.
\begin{align}
\label{eq:linSIR}
\frac{dY}{dt}& = \beta I\\
\frac{dI}{dt}& = (\beta - \alpha) I = \alpha (\phi(t) - 1) I,
\end{align}
and where the number of recovered, $R(t)$, is under this linearization decoupled from the other variables. They further assume the number of deceased is proportional to the number of cumulative infected, offset by a time lag, which allows us to use the same set of linearized equations to model the number of cumulative people deceased without loss of generality. And since $I$ is a function of $Y$, we can combine Eq. \ref{eq:linSIR} via an instantaneous relative growth rate, $\gamma(t) = dY(t) /(Ydt) = \beta I / Y$, 
\begin{equation}
\frac{dY}{dt} = \gamma(t) Y(t) \label{eq:rypdalODE}.
\end{equation}
Due to this linearization, the infection/recovery rate ratio, $\phi(t)$, and the relative growth rate, $\gamma(t)$, will have to change as a function of time to accommodate for the boundary conditions. The latter is typically also parameterized by a scaling factor, $\theta$, representing the shape of the growth,
\begin{equation}
\label{eq:rypdalGamma}
\gamma(t) = \frac{\gamma_{\infty}}{\theta}\left[1 - \left(\frac{Y}{Y_{\infty}}\right)^{\theta} \right] ,
\end{equation}
where $\gamma_{\infty} = \gamma(t\rightarrow \infty)$. This parametrization is the commonly used Richards growth curve \cite{richards1959flexible}, also called $\theta$-logistic growth, and has been used by others \cite{wu2020generalized}. Although not immediately justified in the communicable disease theory, one could imagine that $\theta$ represents non-linear network behavior \cite{petroni2020logistic}. Note that at $\theta=1$, the traditional logistic growth curve is obtained, while at $\theta\rightarrow \infty$ we recover the exponential (Malthusian) explosion.

The observed Gompertzian mortality curves are realized in the limit $\theta \rightarrow 0$, with the relative growth rate, 
\begin{equation}
\label{eq:rypdalLimit}
\gamma(t) = \lim_{\theta \rightarrow 0}\frac{\gamma_{\infty}}{\theta}\left[1 - \left(\frac{Y}{Y_{\infty}}\right)^{\theta} \right]
= \gamma_{\infty}\ln{\frac{Y_{\infty}}{Y}}
\end{equation}

@@

\subsection*{The extended SIR model family is almost Gompertzian}
Without the linearization and the $\theta$-parametrization of the Richards model, one can still almost obtain Gompertzian growth from the communicable disease SIR-like models. Under a double-log-transform of Gompertzian growth Eq. \ref{eq:gomp_solution}, the growth curve is a simple straight line which is one of its characteristic traits \cite{Levitt2020},
\begin{equation}
\label{eq:GOMP_D}
\ln{(\ln{(Y_{\infty}/Y)})} = -\beta t + k,
\end{equation}
with $k=\ln{\ln{\frac{Y_{\infty}}{Y_{0}}}}$. To show this discrepancy between the SIR-like models and the Gompertz model, we follow Carletti et al. (2020) \cite{carletti2020covid} and consider the extended version of the SIR model by including a group of deceased, $D(t)$, such that $N = S(t) + I(t) + R(t) + D(t)$, also called the SIRD model. This extension requires an addition to our original set of equations \ref{eq:SIR}, with an equation for the deceased group's growth at some rate $\delta$ relative to the current infected group,
\begin{equation}
\frac{dD}{dt} = \delta I.
\end{equation}

@@

Without this approximation, the full 3-parameters in Eq. \ref{eq:SIRD_D} of the deceased group can show a curve quite close to a straight line in the double-log-domain even in the initial observations, although there will always be a non-zero concavity (Fig. \ref{fig:Gompertz_vs_SIRD_lnln_infty} and the SI Appendix).
Meanwhile, even though the Gompertz model can be fit with a 3-parameter model as shown in Eq. \ref{eq:GompertzODE}, it can also be simplified to a 2-parameter model estimated through linear regression of log-mortality \cite{Levitt2020},
\begin{equation}
\label{eq:GompertzODE2param}
\frac{d}{dt}\ln{Y(t)} = -\beta\left[\ln{Y(t)} - \ln {Y_{\infty}}\right].
\end{equation}
Thus, through linear regression estimates, the Gompertz model mitigates the possibility of non-identifiability issues of the parameters, which plagues more complex models \cite{roda2020difficult}.

In summary, we see that the SIR model family both lacks the observed immediate decrease in relative growth rate, and makes strong assumptions about the disease spreading network structure that imply non-local interactions. We are thus prompted to look for another model which can explain the observed mortality patterns. For this challenge, we shall first present a stochastic approach and then present a deterministic approach which unifies the SIR family models with the Gompetz curve.

@@

To obtain the deterministic observable, we first take the average in the log-domain (Eq. \ref{eq:microscopic}), 
\begin{equation}
d \;\mathbb E[\ln{F(t)}] = -\beta \mathbb E[\ln{F(t)}] dt,
\end{equation}
where $\mathbb E$ signifies the averaging operation.
Then we use the property that for the log-normal quantity $F(t)$, the average of its logarithm is the logarithm of the \emph{median} quantity \cite{petroni2020gompertz}, where we denote the median of $F(t)$ as $M(t)$, 
\begin{equation}
\label{eq:medianGomp}
d\ln{M(t)} = -\beta \ln{M(t)} dt,
\end{equation}
which corresponds with the familiar deterministic Gompertz differential equation in Eq. \ref{eq:GompertzODE} with $Y(t) = Y_\infty M(t)$. 
Comparing the stochastic stressor $\sigma$ in Eq. \ref{eq:microscopicOriginal} with the deterministic growth equation in Eq. \ref{eq:GompertzODE} gives $\nu=\sigma/2$, suggesting that the stressor is indeed the source of growth, while $\beta$ is the growth-limiting factor. This comparison is clear after taking the median of Eq. \ref{eq:microscopicOriginal} wherein the cross-term of $F(t)dW(t)$ goes to zero due to the independence of $F(t)$ and $dW(t)$.
We also see this by comparing equations \ref{eq:rypdalODE} and \ref{eq:rypdalLimit} with the stochastic counterpart in Eq. \ref{eq:microscopicOriginal} that the final growth level is governed by the stressor magnitude,
\begin{equation}
\sigma = 2\gamma_{\infty}\ln{Y_{\infty}}.
\end{equation}

@@

\begin{equation}
\frac{d x_i}{dt} = -\beta x_i\sum_j{a_{ij}}\ln{(1 - s_j)} - \alpha = -\beta x_i\sum_j{a_{ij}}\ln{x_j} - \alpha x_i \quad \forall i,
\end{equation}
As we are interested in the aggregate macroscopic behavior, we take averages in the log-domain, exploit our setup where $\mathbf{A}$ is a single mapping from one node to another, and simplify to
\begin{equation}
\frac{d}{{dt}} \ln\left[\prod_i^N{x_i}\right]^{\frac{1}{N}} = -\beta\ln\left[\prod_i^N{x_i}\right]^{\frac{1}{N}} - \frac{\alpha}{N}.
\end{equation}
If $\alpha\rightarrow 0$, then this equation will exhibit Gompertzian growth in the geometric ensemble average of microscopic units. 
An interpretation of $\alpha$ approaching $0$ could be that the limiting factor emerges purely from the growth rate without the need for a second growth-limiting parameter, in line with the previous crucial hypothesis of causality reversal stated above. One further simplification could be seen in equating the logarithm of the geometric mean with the logarithm of the median of the set of $x_i$ to obtain Eq. \ref{eq:medianGomp},
\begin{equation}
\frac{d}{dt}\ln{M(t)} = -\beta \ln{M(t)}.
\end{equation}
We thus see that in a mathematically well-defined scenario, nodes in the network respond instantaneously and are correlated with each other as under a global biosphere disturbance.

@@

\section*{Author Contributions}
M. A. Haugen: Conceptualization, Methodology, Software, Writing. D. Gilbert: Conceptualization, Reviewing and Editing.

\section*{Appendix}
A elaboration of the Gompertz equation in \ref{eq:GompertzODEsimple} is given below, which includes an estimate of the abundance as time goes to infinity,
\begin{equation}
\label{eq:GompertzODE}
\frac{d}{dt}\ln{Y(t)} = -\beta\ln{\frac{Y(t)}{\tilde{Y}}} + \nu,
\end{equation} 
% Redundant definitions for full script
with constants $\tilde{Y}$, $\beta$, and $\nu$, and whose solution is given as
\begin{equation}
\label{eq:gomp_solution}
Y(t) = Y_\infty \left(\frac{Y_0}{Y_\infty}\right)^{e^{-\beta t}},
\end{equation}
where $Y_{0}=Y(t = 0)$ and $Y_{\infty}=Y(t\rightarrow \infty)=\tilde{Y}e^{\nu/\beta}$.

>>>>

% \author{Matz A. Haugen, \\
% Independent Reseacher, Oslo\\
% \texttt{matzhaugen@gmail.com} \\
%   \And
%   Dorothea Gilbert \\
%   Independent Reseacher, Oslo\\
% }

@@

A line of reasoning employed by Rypdal and Rypdal (2020) \cite{Rypdal2020} to obtain the Gompertz curve is to linearize the SIR model by assuming
both the number of infected, $I(t)$, and cumulative infected, $Y(t)$, are much less than the initial pool of susceptible people, $N \gg Y \ge I$, which seems well founded at the national level, viz.
\begin{align}
\label{eq:linSIR}
\frac{dY}{dt}& = \beta I\\
\frac{dI}{dt}& = (\beta - \alpha) I = \alpha (\phi(t) - 1) I,
\end{align}
and where the number of recovered, $R(t)$, is under this linearization decoupled from the other variables. They further assume the number of deceased is proportional to the number of cumulative infected, offset by a time lag, which allows us to use the same set of linearized equations to model the number of cumulative people deceased without loss of generality. And since $I$ is a function of $Y$, we can combine Eq. \ref{eq:linSIR} via an instantaneous relative growth rate, $\gamma(t) = dY(t) /(Ydt) = \beta I / Y$, 
\begin{equation}
\frac{dY}{dt} = \gamma(t) Y(t) \label{eq:rypdalODE}.
\end{equation}
Due to this linearization, the infection/recovery rate ratio, $\phi(t)$, and the relative growth rate, $\gamma(t)$, will have to change as a function of time to accommodate for the boundary conditions. The latter is typically also parameterized by a scaling factor, $\theta$, representing the shape of the growth,
\begin{equation}
\label{eq:rypdalGamma}
\gamma(t) = \frac{\beta}{\theta}\left[1 - \left(\frac{Y}{Y_{\infty}}\right)^{\theta} \right].
\end{equation}
This parametrization is the commonly used Richards growth curve \cite{richards1959flexible}, also called $\theta$-logistic growth, and has been used by others \cite{wu2020generalized}. Although not immediately justified in the communicable disease theory, one could imagine that $\theta$ represents non-linear network behavior \cite{petroni2020logistic}. Note that at $\theta=1$, the traditional logistic growth curve is obtained, while at $\theta\rightarrow \infty$ we recover the exponential (Malthusian) explosion.

The observed Gompertzian mortality curves are realized in the limit $\theta \rightarrow 0$, with the relative growth rate, 
\begin{equation}
\label{eq:rypdalLimit}
\gamma(t) = \lim_{\theta \rightarrow 0}\frac{\beta}{\theta}\left[1 - \left(\frac{Y}{Y_{\infty}}\right)^{\theta} \right]
= \beta\ln{\frac{Y_{\infty}}{Y}},
\end{equation}
such that 
\begin{equation}
\frac{d \ln Y}{dt} = -\beta \ln \frac{Y}{Y_\infty} \label{eq:gompertzODEfirst}.
\end{equation}

@@

\subsection*{The extended SIR model family is almost Gompertzian}
Without the linearization and the $\theta$-parametrization of the Richards model, one can still almost obtain Gompertzian growth from the communicable disease SIR-like models. Under a double-log-transform of Gompertzian growth, the growth curve is a simple straight line which is one of its characteristic traits \cite{Levitt2020},
\begin{equation}
\label{eq:GOMP_D}
\ln{(\ln{(Y_{\infty}/Y)})} = -\beta t + k,
\end{equation}
with $k=\ln{\ln{\frac{Y_{\infty}}{Y_{0}}}}$. To show this discrepancy between the SIR-like models and the Gompertz model, we follow Carletti et al. (2020) \cite{carletti2020covid} and consider the extended version of the SIR model by including a group of deceased, $D(t)$, such that $N = S(t) + I(t) + R(t) + D(t)$, also called the SIRD model. This extension requires an addition to our original set of equations \ref{eq:SIR}, with an equation for the deceased group's growth at some rate $\delta$ relative to the current infected group,
\begin{equation}
\frac{dD}{dt} = \delta I.
\end{equation}

@@

Without this approximation, the full 3-parameters in Eq. \ref{eq:SIRD_D} of the deceased group can show a curve quite close to a straight line in the double-log-domain even in the initial observations, although there will always be a non-zero concavity (Fig. \ref{fig:Gompertz_vs_SIRD_lnln_infty} and the SI Appendix).

Meanwhile, even though the Gompertz model can be fit with a 3-parameter model, it can also be simplified to a 2-parameter model estimated through linear regression of log-mortality \cite{Levitt2020},
\begin{equation}
\label{eq:GompertzODE2param}
\frac{d}{dt}\ln{Y(t)} = -\beta\left[\ln{Y(t)} - \ln {Y_{\infty}}\right].
\end{equation}

The 3-parameter model has one extra degree of freedom which can stabilize the asymptotic abundance estimate, 
\begin{equation}
\label{eq:GompertzODE}
\frac{d}{dt}\ln{Y(t)} = -\beta\ln{\frac{Y(t)}{\tilde{Y}}} + \nu,
\end{equation} 
% Redundant definitions for full script
with constants $\tilde{Y}$, $\beta$, and $\nu$. The Gompertz growth solution is given as
\begin{equation}
\label{eq:gomp_solution}
Y(t) = Y_\infty \left(\frac{Y_0}{Y_\infty}\right)^{e^{-\beta t}},
\end{equation}
where $Y_{0}=Y(t = 0)$ and $Y_{\infty}=\tilde{Y}e^{\nu/\beta}$.

Whether though the 3-parameter model or the 2-parameter linear model, the Gompertz model is relatively parsimonious in its parameters and thus mitigates the possibility of non-identifiability issues, which plagues more complex models \cite{roda2020difficult}.

In summary, we see that the SIR model family both lacks the observed immediate decrease in relative growth rate, and makes strong assumptions about the disease spreading network structure that imply non-local interactions. We are thus prompted to look for another model which can explain the observed mortality patterns. For this challenge, we shall first present a stochastic approach and then present a deterministic approach which unifies the SIR family models with the Gompertz curve.

@@

To obtain the deterministic observable, we first take the average in the log-domain (Eq. \ref{eq:microscopic}), 
\begin{equation}
d \;\mathbb E[\ln{F(t)}] = -\beta \mathbb E[\ln{F(t)}] dt,
\end{equation}
where $\mathbb E$ signifies the averaging operation.
Then we use the property that for the log-normal quantity $F(t)$, the average of its logarithm is the logarithm of the \emph{median} quantity \cite{petroni2020gompertz}, where we denote the median of $F(t)$ as $M(t)$, 
\begin{equation}
\label{eq:medianGomp}
d\ln{M(t)} = -\beta \ln{M(t)} dt,
\end{equation}
which corresponds with the familiar deterministic Gompertz differential equation in Eq. \ref{eq:GompertzODE} with $Y(t) = Y_\infty M(t)$. 
Comparing the stochastic stressor $\sigma$ in Eq. \ref{eq:microscopicOriginal} with the deterministic growth equation in Eq. \ref{eq:GompertzODE} gives $\nu=\sigma/2$, suggesting that the stressor is indeed the source of growth, while $\beta$ is the growth-limiting factor. This comparison is clear after taking the median of Eq. \ref{eq:microscopicOriginal} wherein the cross-term of $F(t)dW(t)$ goes to zero due to the independence of $F(t)$ and $dW(t)$.

@@

\begin{equation}
\frac{d x_i}{dt} = -\beta x_i\sum_j{a_{ij}}\ln{(1 - s_j)} - \alpha x_i = -\beta x_i\sum_j{a_{ij}}\ln{x_j} - \alpha x_i \quad \forall i,
\end{equation}
As we are interested in the aggregate macroscopic behavior, we take averages in the log-domain, exploit our setup where $\mathbf{A}$ is a single mapping from one node to another, and simplify to
\begin{equation}
\frac{d}{{dt}} \ln\left[\prod_i^N{x_i}\right]^{\frac{1}{N}} = -\beta\ln\left[\prod_i^N{x_i}\right]^{\frac{1}{N}} - \frac{\alpha}{N}.
\end{equation}
If $\alpha\rightarrow 0$, then this equation will exhibit Gompertzian growth in the geometric ensemble average of microscopic units. 
An interpretation of $\alpha$ approaching $0$ could be that the limiting factor emerges purely from the growth rate without the need for a second growth-limiting parameter, in line with the previous crucial hypothesis of causality reversal stated above. One further simplification could be seen in equating the logarithm of the geometric mean with the logarithm of the median of the set of $x_i$ to obtain Eq. \ref{eq:medianGomp},
\begin{equation}
\frac{d}{dt}\ln{M(t)} = -\beta \ln{M(t)}.
\end{equation}
We thus see that in a mathematically well-defined scenario, nodes in the network respond instantaneously and are correlated with each other as under a global biosphere disturbance.

@@

% \section*{Author Contributions}
% M. A. Haugen: Conceptualization, Methodology, Software, Writing. D. Gilbert: Conceptualization, Reviewing and Editing.

### 2023-04-04 letterToEditor/letter.tex
respond to comments2

\recipient{To}{Editor, \\
               Journal of Mathematical Biology,\\ 
               }
\date{\today}
\opening{\textbf{Initial submission of manuscript titled: ``Unifying the communicable disease spreading paradigm with Gompertzian growth''}}
\closing{Sincerely yours, \vspace{-1em}}

@@

If you do decide to review this paper you may find the enclosed list of possible referees helpful. Some of them are listed in the references of the manuscript. Due to the conclusions of our paper, I request that this paper receives more reviews than your average, and at least 3 reviews.

@@

% \item Nigel D. Goldenfeld 
% \item Evans, Steven N.
% \item Altshuler, Boris L.
% \end{enumerate}
\textbf{List of possible referees}
\begin{enumerate}
\item Tom Britton, Dean of Mathematics and Physics, Stockholm University, tom.britton@math.su.se
\item Mark Lewis, marklewis@uvic.ca, Department of Mathematics and Statistics, University of Victoria
\item Alexandru Hening, Department of Mathematics at Texas A and M University,  ahening@tamu.edu
\item K-Y. Lam, Mathematics,  the Ohio State University, Columbus, lam.184@osu.edu 
\item X. Zou, Western University, London, Ontario, xzou@uwo.ca
\item S. Petrovskii, Mathematical Sciences, University of Leicester, sp237@le.ac.uk
\item Pierre Magal, Institut de Mathématiques de Bordeaux, Université de Bordeaux, pierre.magal@u-bordeaux.fr 
\item Vladimir N. Binhi, vnbin@mail.ru, Prokhorov General Physics Institute of the Russian Academy of Sciences, 38 Vavilov St., 119991 Moscow, Russia
Is familiar with the biological effects due geomagnetic disturbances and fields.
\item Andrei B. Rubin, rubin@biophysics.msu.ru, Faculty of Biology, Lomonosov Moscow State University, Leninskie Gory 1/12, 119234 Moscow, Russia
Is familiar with the biological effects due geomagnetic disturbances and fields.
\item Viacheslav V. Krylov, I.D. Papanin Institute for Biology of Inland Waters Russian Academy of Sciences, Borok, Russian Federation, krylovviacheslav@mail.ru
Is intimately familiar with geomagnetic disturbances' effect on biology.
\item Marcin Molski, mamolski@amu.edu.pl, Adam Mickiewicz University, physics, Gompertz and quantum systems. Has done work on connecting the gompertz model to quantum coherent systems.
\item Francesco Piazza, Francesco.Piazza@cnrs-orleans.fr, Max Planck Institute for the Physics of Complex Systems, physics, SIR models.
Has written one of the central papers modeling covid disease propagation from the SIR perspective. We have based our methods on their paper.
\item Robert M May, robert.may@zoo.ox.ac.uk, Professor of Zoology, University of Oxford.
\end{enumerate}

>>>>

\recipient{To}{Editor, \\
               Journal of Physics A: Mathematical and Theoretical,\\ 
               }
\date{\today}
\opening{\textbf{Initial submission of manuscript titled: ``Unifying the communicable disease spreading paradigm with Gompertzian growth''}}
\closing{Sincerely yours, \vspace{-1em}}

@@

Due to the conclusions of our paper, I request that this paper receives more reviews than your average, and at least 3 reviews. Even though we have recommended a few reviewers, perhaps you know someone who is better qualified to review, especially since we are submitting to a special section of your journal.

@@

% \item Nigel D. Goldenfeld 
% \item Evans, Steven N.
% \item Altshuler, Boris L.
% \end{enumerate}
% \textbf{List of possible referees}
% \begin{enumerate}
% % IOP REVIEWERS
% \item Martin Evans,Professor of Statistical Physics,  https://www2.ph.ed.ac.uk/~mevans/, m.evans@ed.ac.uk
% \item Prof Ralf Metzler, Institute of Physics & Astronomy
% University of Potsdam
% Karl-Liebknecht-Str 24/25, Haus 28
% D-14476 Potsdam-Golm 
% \item Michael Victor Berry, U of Bristol, Prof. of Physics, asymptotico(add at bristol.ac.uk) 
% \item Ronald Dickman, Professor of Physics, UFMG, dickman@fisica.ufmg.br
% \item Ewa Gudowska-Nowak, ewa.gudowska-nowak[at]uj.edu.pl , 
% Marian Smoluchowski
% Institute of Physics,
% Jagiellonian University
% \item Shamik Gupta Tata Institute of Fundamental Research, Mumbai, India, shamik.gupta-At-theory.tifr.res.in (Official)
% \item Alain Barrat, CNRS & Aix-Marseille Université, France , Alain.Barrat@cpt.univ-mrs.fr
% \item Daniel ben-Avraham, dbenavra@clarkson.edu, Professor of Physics, Clarkson University
% \item Tom Britton, Dean of Mathematics and Physics, Stockholm University, tom.britton@math.su.se
% \item Mark Lewis, marklewis@uvic.ca, Department of Mathematics and Statistics, University of Victoria
% \item Alexandru Hening, Department of Mathematics at Texas A and M University,  ahening@tamu.edu
% \item K-Y. Lam, Mathematics,  the Ohio State University, Columbus, lam.184@osu.edu 
% \item X. Zou, Western University, London, Ontario, xzou@uwo.ca
% \item S. Petrovskii, Mathematical Sciences, University of Leicester, sp237@le.ac.uk
% \item Pierre Magal, Institut de Mathématiques de Bordeaux, Université de Bordeaux, pierre.magal@u-bordeaux.fr 
% \item Vladimir N. Binhi, vnbin@mail.ru, Prokhorov General Physics Institute of the Russian Academy of Sciences, 38 Vavilov St., 119991 Moscow, Russia
% Is familiar with the biological effects due geomagnetic disturbances and fields.
% \item Andrei B. Rubin, rubin@biophysics.msu.ru, Faculty of Biology, Lomonosov Moscow State University, Leninskie Gory 1/12, 119234 Moscow, Russia
% Is familiar with the biological effects due geomagnetic disturbances and fields.
% \item Viacheslav V. Krylov, I.D. Papanin Institute for Biology of Inland Waters Russian Academy of Sciences, Borok, Russian Federation, krylovviacheslav@mail.ru
% Is intimately familiar with geomagnetic disturbances' effect on biology.
% \item Marcin Molski, mamolski@amu.edu.pl, Adam Mickiewicz University, physics, Gompertz and quantum systems. Has done work on connecting the gompertz model to quantum coherent systems.
% \item Francesco Piazza, Francesco.Piazza@cnrs-orleans.fr, Max Planck Institute for the Physics of Complex Systems, physics, SIR models.
% Has written one of the central papers modeling covid disease propagation from the SIR perspective. We have based our methods on their paper.
% \item Robert M May, robert.may@zoo.ox.ac.uk, Professor of Zoology, University of Oxford.
\end{enumerate}
